### 2016-07-08 Projects/DelCo_2012/Report/DelCo_2012.tex
Craig's paragraph and table + text changes in section 3

To better quantify the effectiveness of the suppression agents (CAF, water), the maximum temperature differences associated with each gas cooling event for each agent from the three thermocouple trees were analyzed. Figure~\ref{fig:combined_all} compares the temperature differences at each thermocouple position (via a scatterplot) for 30 of the 88 experiments, 15 water and 15 CAF, which were directly comparable.  These experiments included solid stream, straight streams, and narrow fog streams at both nozzle positions.  While the total number of experiments was 88, there where a number of experiments that were performed which did not have a directly comparable experiments between water and CAF.  For example a set of 12 Class A solutions experiments were conducted.  Gas cooling experiments were conducted with water applied on a wide fog nozzle setting, no significant differences were noted so a comparable series with wide fog CAF was not conducted.  Similarly larger water flows were examined but 120 gpm was the maximum water flow that could be generated with the CAFS, so those experiments are not included in this comparison.  The Mitchell report does include all of the data for the 88 experiments \cite{Mitchell:1}

\textbf{Craig - the table of 30 could go here and the R squared paragraph could go with the first discussion of the red line below}

@@

@@

There were 12 suppression tests conducted in the single story purpose-built concrete structure. Four of the tests used a wood pallet fuel load (Section~\ref{sec:fire_suppression_pallet_fuel}) and the remaining tests used a fuel load comprised of typical household furnishings that included wooden wall paneling, two sofas, two overstuffed chairs, carpet, and padding (Section~\ref{sec:fire_suppression_furniture_fuel}). Ignition occurred in the rear of the single story structures and the available ventilation was through the front doorway and for several tests a rear window was opened to provide additional ventilation (see Table~\ref{tab:Test_Descriptions}). The fires were intended to replicate room and contents fires typically seen within residential structures. Hose lines were pre-deployed with fixed nozzles set in two locations: one in the hallway on the approach to the burn room (hallway nozzle) and one just inside the burn room to the right side of the hallway (room nozzle). Each nozzle flowed 15~s of water or compressed air foam depending on the test. The hallway nozzle was intended to simulate a crew advancing into a structure towards the burn room. Suppression from this location would be in the form of gas cooling as the stream was not aimed at the seat of the fire (indirect interior attack). The room nozzle was intended to simulate a crew which had advanced into a structure, entered the fire room, positioned just to the side of the entry door, and began suppression on the seat of the fire (direct interior attack). Table~\ref{tab:Test_Descriptions} provides an overview of the suppression tests conducted.

@@

Table~\ref{tab:Test_Results} provides the impact of the two nozzle applications for the water and CAF tests. The table provides change in temperature 0.3~m (1~ft) below the ceiling and 0.3~m (1~ft) above the floor as well as change in heat flux 0.15~m (6~in) above the floor at the measurement location closest to the fire. These values reported in the table are immediately before and after the 15~s application period to assess the application agent, stream, and location on interior conditions.

@@

\subsection{Iowa State Univeristy Fire Flow Formula (1963)~\cite{NFPA}}
The Iowa Formula for determining the required fire flowrate for a given structure was developed, by Keith Royer and Bill Nelson, based on theoretical heat absorption capabilities of water. The required fire flowrate in gpm is equal to the volume of the fire compartment (in cubic feet) divided by 100:
\begin{equation}\label{eq:isu_form}
\required fire flowrate = Volume / 100
\end{equation}
Due to inefficiencies in applying the water, some feel that the fire flow value from the Iowa State University formula should be multiplied by a factor of two to four~\cite{NFPA}. Applying the  based on floor area yields an application density of 3.3~Lpm/m$^2$ (0.08~gpm/ft$^2$) for ceiling heights of 2.44~m (8~ft) or 4.1~Lpm/m$^2$ (0.10~gpm/ft$^2$) for ceiling heights of 3.05~m (10~ft). Our fire rooms ranged in floor area from approximately 66~m$^3$ (2330~ft$^3$) to 125~m$^3$ (4410~ft$^3$).  Therefore our required flowrate for the fire room in the single story structure based on the Iowa State University formula would be approximately 87~Lpm (23~gpm).  The required flowrate for the fire room in the lower level of the two story would be approximately 168~Lpm (44~gpm).

\subsection{National Fire Academy (1884)~\cite{Klaene:1}}
The National Fire Academy (NFA) developed a formula based on a study of fire flows, which were successful in controlling a large number of working fires~\cite{Klaene:1}. This formula sets the required fire flow, in gpm, equal to the floor area (in square feet) divided by 3.
\begin{equation}\label{eq:nfa_form}
\required fire flowrate = Area / 3
\end{equation}

>>>>

To better quantify the effectiveness of the suppression agents (CAF, water), the maximum temperature differences associated with each gas cooling event for each agent from the three thermocouple trees were analyzed. Figure~\ref{fig:combined_all} compares the temperature differences at each thermocouple position (via a scatterplot) for 30 of the 88 experiments.  The 30 experiments are actually pairs of experiments, 15 water and 15 CAF, which were directly comparable, see Table~\ref{tab:Gas_Cooling_Descriptions}.  These experiments included solid stream, straight streams, and narrow fog streams at both nozzle positions.  While the total number of experiments was 88, there where a number of experiments that were performed which did not have a directly comparable experiments between water and CAF.  For example a set of 12 Class A solutions experiments were conducted.  Gas cooling experiments were conducted with water applied on a wide fog nozzle setting, no significant differences were noted so a comparable series with wide fog CAF was not conducted.  Similarly larger water flows were examined but 120 gpm was the maximum water flow that could be generated with the CAFS, so those experiments are not included in this comparison.  The Mitchell report does include all of the data for the 88 experiments \cite{Mitchell:1}

\begin{table}[!ht]
\centering
\caption{Gas Cooling Comparison Test Descriptions}\label{tab:Gas_Cooling_Descriptions}
\begin{tabular}{clll}
\toprule[1.5pt]
Water Test $\#$  & CAF Test  $\#$	& Stream			  & Position  \\
\midrule
 2               & 6                &  Solid Stream       &  Middle   \\
 3               & 7                &  Solid Stream       &  Middle   \\
 4               & 10               &  Solid Stream       &  Middle   \\
 9               & 11               &  Solid Stream       &  Middle   \\
 17              & 52               &  Solid Stream       &  Middle   \\
 18              & 53               &  Solid Stream       &  Middle   \\
 19              & 54               &  Solid Stream       &  Middle   \\
 16              & 45               &  Straight Stream    &  Middle   \\
 20              & 46               &  Straight Stream    &  Middle   \\
 21              & 47               &  Straight Stream    &  Middle   \\
 22              & 48               &  Straight Stream    &  Middle   \\
 41              & 50               &  Straight Stream    &  Middle   \\
 29              & 12               &  Straight Stream    &  Full Back   \\
 33              & 13               &  30$^{\circ}$ Fog   &  Full Back   \\
 34              & 14               &  30$^{\circ}$ Fog   &  Full Back   \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

The quality of the fit of the red dashed line to the scatter in gas cooling data is represented by the coefficient of determination (R$^2$ on Fig.~\ref{fig:combined_all}). The coefficient of determination provides a way to quantify how significant the variation of the data is compared the the least squared error line. Here, the higher the value of R$^2$, the better the ordinary least square line describes the relationship between water and CAF temperature difference comparisons.

@@

There were 12 suppression tests conducted in the single story purpose-built concrete structure. Eight of the experiments used a fuel load comprised of typical household furnishings that included wooden wall paneling, two sofas, two overstuffed chairs, carpet, and padding (Section~\ref{sec:fire_suppression_furniture_fuel}). The other four experiments used wood pallet fuel load in a space with the ceiling and walls composed of wood based materials (Section~\ref{sec:fire_suppression_pallet_fuel}). Ignition occurred in the rear of the single story structures and the available ventilation was through the front doorway and for eight of the experiments a rear window was "opened" to provide additional ventilation (see Table~\ref{tab:Test_Descriptions}). The fires were intended to replicate post-flashover residential room and contents fires. Hose lines were pre-deployed with fixed nozzles set in two locations: one in the hallway on the approach to the burn room (hallway nozzle) and one just inside the burn room to the right side of the hallway (room nozzle). Each nozzle flowed 15~s of water or compressed air foam depending on the test. The hallway nozzle was intended to simulate a crew advancing into a structure towards the burn room. Suppression from this location would be in the form of gas cooling as the stream was not aimed at the seat of the fire (indirect interior attack).  After the hallway nozzle was shutdown, there was a 15~s time period to simulate the crew advancing into the doorway of the fire room. The room nozzle was positioned just to the side of the entry door.  It was also opened and flowed for approximately 15~s. The room nozzles was aimed in the direction of the seat of the fire (direct interior attack). Table~\ref{tab:Test_Descriptions} provides an overview of the suppression tests conducted.

@@

Table~\ref{tab:Test_Results} provides the impact of the two nozzle applications for the water and CAF tests. The table provides change in temperature 0.3~m (1~ft) below the ceiling and 0.3~m (1~ft) above the floor as well as change in heat flux 0.15~m (6~in) above the floor at the measurement location closest to the fire. These values reported in the table are immediately before and after the 15~s application period to assess the application agent, stream, and location on interior conditions.  Remeber that there was a 15~s gap between the hallway nozzle flow and the fire room nozzle flow.  As a result you can see the amount of temperature change during the period in between nozzle flows by comparing the temperature at the end of the hallway flow period with the temperature at the beginning of the room flow period.

@@

\subsection{Iowa State Univeristy Fire Flow Formula (1963)~\cite{NFPA}}
The Iowa Formula for determining the required fire flowrate for a given structure was developed, by Keith Royer and Bill Nelson, based on theoretical heat absorption capabilities of water. The required fire flowrate in gpm is equal to the volume of the fire compartment (in cubic feet) divided by 100:
\begin{equation}\label{eq:isu_form}
required fire flowrate = Volume / 100
\end{equation}
Due to inefficiencies in applying the water, some feel that the fire flow value from the Iowa State University formula should be multiplied by a factor of two to four~\cite{NFPA}. Applying the  based on floor area yields an application density of 3.3~Lpm/m$^2$ (0.08~gpm/ft$^2$) for ceiling heights of 2.44~m (8~ft) or 4.1~Lpm/m$^2$ (0.10~gpm/ft$^2$) for ceiling heights of 3.05~m (10~ft). Our fire rooms ranged in floor area from approximately 66~m$^3$ (2330~ft$^3$) to 125~m$^3$ (4410~ft$^3$).  Therefore our required flowrate for the fire room in the single story structure based on the Iowa State University formula would be approximately 87~Lpm (23~gpm).  The required flowrate for the fire room in the lower level of the two story would be approximately 168~Lpm (44~gpm).

\subsection{National Fire Academy (1884)~\cite{Klaene:1}}
The National Fire Academy (NFA) developed a formula based on a study of fire flows, which were successful in controlling a large number of working fires~\cite{Klaene:1}. This formula sets the required fire flow, in gpm, equal to the floor area (in square feet) divided by 3.
\begin{equation}\label{eq:nfa_form}
required fire flowrate = Area / 3
\end{equation}
